---
output: 
  html_document:
    fig_caption: true
editor_options: 
  chunk_output_type: console
  markdown:
    wrap: 80
---

# (PART) Data downloads and decisions {.unnumbered}

```{r library-ch1, include=FALSE}
library(sf)
library(feather)
library(GGally)
library(mapview)
library(ggthemes)
library(tidyverse)

```

# Download and Merge Aquasat data

[[I've removed all `eval=F` and `cache=T` so that the .html will open and show
all figs correctly. My download of files didn't take more than a few minutes, so
I'm not sure if that's an internet thing? Or maybe because my machine is
relatively fast? I'm happy to put back in on next round.]]

## AquaSat download

AquaSat is held in a figshare collection with many pieces. For this work we will
need most of the data held there, which we will access through direct download.
The URLs are just from the AquaSat collection links.

<details>

<summary>Code to download data.</summary>

```{r download-data}
# check to see if the data have been downloaded. we assume that if the directory exists, data have been downloaded.

download_dir <- "../data/in/"

if(!dir.exists(download_dir)) {
  # if it doesn't, make the folder and download AquaSat files
  dir.create(download_dir, recursive = TRUE)
  ## AquaSat Download data
  # In-situ raw data with methods
  download.file('https://figshare.com/ndownloader/files/15475154',
                mode = 'wb',# Needs to be written in binary for some reason 
                destfile = file.path(download_dir, 'aq_situ.zip'))

    unzip(file.path(download_dir, 'aq_situ.zip'),
          exdir = file.path(download_dir, 'aq_situ'))
  
  # Site Inventory with type, because it's not in the other inventory
  # Stupid aquasat developer (me/mattross!)
  download.file('https://figshare.com/ndownloader/files/24720434',
                mode = 'wb',
                destfile = file.path(download_dir, 'inv.feather'))
  
  # Unique site inventory 
    download.file('https://figshare.com/ndownloader/files/24720437',
                mode = 'wb',
                destfile = file.path(download_dir, 'unq_site.feather'))
  
  ## Ecoregion data
  download.file('https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/cec_na/na_cec_eco_l2.zip', 
                destfile = file.path(download_dir, 'eco2.zip'))
    unzip(file.path(download_dir, 'eco2.zip'),
          exdir = file.path(download_dir, 'ecoregion'))

}

```

</details>

## AquaSat Read and join.

[[note here: 'data/in/clouds.feather' nor WRS2 shapefile
'data/in/wrs/WRS2_descending.shp' are parts of the downloaded files, commented
out for rendering purposes]]

<details>

<summary>Code to collate downloaded data.</summary>

```{r read-join-aquasat}

data_out_dir <- "../data/out/"

# this chunk only needs to be run if the out directory doesn't exist
if(!dir.exists(data_out_dir)) {
  # grab monitoring location identifier (SiteID/MonitorongLocationIdentifier) and
  # short-hand site type (river/lake/estuary/facilty)
  inv_type <- read_feather(file.path(download_dir, 'inv.feather')) %>%
    select(SiteID = MonitoringLocationIdentifier,
           type = ResolvedMonitoringLocationTypeName) %>%
    mutate(type = if_else(grepl('Lake',type),'Lake',type)) %>% 
    #drop 'Facility' types
    filter(type != 'Facility')
  
  # combine with site lat/long from unique site inventory
  site_vis <- read_feather(file.path(download_dir, 'unq_site.feather')) %>%
    inner_join(inv_type) %>%
    distinct(SiteID,lat,long,type)
  
  # load complete in-situ data, 6.5 million records. 
  in_vis <- read_csv(file.path(download_dir, 'aq_situ/in-situ/wqp_lagos_unity.csv')) 
  
  # # Get clouds and path row sites
  # clouds <- read_feather('data/in/clouds.feather') %>%
  #   mutate(date = as.Date(SENSING_TIME)) %>%
  #   select(PATH = WRS_PATH, 
  #          ROW = WRS_ROW,
  #          clouds = CLOUD_COVER,
  #          date)
  # 
  # site_path_row <- site_vis %>%
  #   st_as_sf(.,coords = c('long','lat'), crs = 4326) %>%
  #   st_join(.,st_read('data/in/wrs/WRS2_descending.shp')) %>%
  #   select(SiteID, PATH, ROW) %>%
  #   as_tibble(.) %>%
  #   select(-geometry)

}

```

</details>

## Data Evalualtion and Subset

### Selecting only simultaneous observations

For the purposes of this analysis, we only care about observations that have
complete simultaneous observations across the four parameters of interest:
chlorophyll *a*, DOC, TSS, and Secchi disk depth.

<details>

<summary>Code to subset for simultaneous observations.</summary>

```{r simultaneous}
if(!dir.exists(data_out_dir)) {
  #create directory path
  dir.create(data_out_dir, recursive = TRUE)
  simul_vis <- in_vis %>%
    select(-p_sand) %>%
    filter(if_all(c(chl_a,doc,tss,secchi), ~!is.na(.))) %>% 
    inner_join(site_vis) %>%
    #Set some reasonable thresholds, AquaSat is too generous
    filter(secchi < 15,
           chl_a < 1000,## ug/L
           tss < 1000, ## mg/L
           doc < 50)
  
  no_secchi <- in_vis %>%
    select(-p_sand) %>%
    filter(if_all(c(chl_a,doc,tss), ~!is.na(.))) %>% 
    inner_join(site_vis) %>%
    #Set some reasonable thresholds, AquaSat is too generous
    filter(
      chl_a < 1000,## ug/L
      tss < 1000, ## mg/L
      doc < 50)
  
  # For speed
  write_feather(simul_vis, file.path(data_out_dir, 'simul.feather'))
  #For reproducibility
  write_csv(simul_vis, file.path(data_out_dir, 'simul.csv'))
  #For speed
  write_feather(no_secchi, file.path(data_out_dir, 'no_secchi.feather'))
}

```

</details>

## Where are sites with simultaneous observations of clarity constituents?

<details>

<summary>Code to create map figure.</summary>

```{r make-map-simultaneous}

simul_vis <- read_csv(file.path(data_out_dir, 'simul.csv'))

unique_simul <- simul_vis %>%
  distinct(SiteID, lat, long, type) %>%
  st_as_sf(.,coords = c('long','lat'), crs = 4326)

mapviewOptions(fgb = FALSE) # remove flat geobuff which breaks display
simul_map <- mapview(unique_simul, zcol = 'type')

```

</details>

```{r map-simultaneous, echo=FALSE, fig.cap= "Map of all sites with simultaneous chlorohpyll *a*, DOC, TSS, and Secchi disk depth measurements within the Aquasat data set."}
#as a note, fig.cap will not work for this chunk, because markdown does not see interactive mapview as a figure.
simul_map

```

## What is the general relationship between variables in log-log space?

<details>

<summary>Code to create scatter plot matrix figure.</summary>

```{r four-vars-scatter}
log_simul <- simul_vis %>%
  mutate(across(c(secchi,chl_a,tss,doc,tis), log10)) %>%
  filter(if_all(c(chl_a,doc,secchi,tss), ~!is.na(.) & . < Inf & . > -Inf)) 

vars_splom <- log_simul %>%
  # randomly sample 20% of each site type to display info.
  sample_frac(0.2) %>%
  ungroup() %>%
  select(secchi,chl_a,tss,doc,type) %>%
  ggpairs(lower = list(continuous = wrap('points',shape = 1)),
          diag = list(continuous = wrap('densityDiag', alpha = 0.5)),
          mapping = ggplot2::aes(color = type),
          columns = c('secchi','chl_a','tss','doc')) +
  theme_few() + 
  scale_color_manual(values = c('seagreen3','skyblue3','saddlebrown'))

```

</details>

```{r splom-fig, echo=F, fig.cap="Scatter plot matrix of four parameters for a random subsample of data weighted by water type."}

vars_splom

```
