---
output: 
  html_document:
    fig_caption: true
editor_options: 
  chunk_output_type: console
  markdown:
    wrap: 80
---

# Data quality assessments

```{r load-pkgs-ch2, include = F}
library(sf)
library(feather)
library(GGally)
library(mapview)
library(ggpmisc)
library(Metrics)
library(kableExtra)
library(broom)
library(ggthemes)
library(knitr)
library(tidyverse)

opts_chunk$set(warning = F, message = F)

data_out_dir <- "data/out"

```

## Chl-a to biomass

While it is more common in coastal and oceanic research, inland water scientists
rarely measure water clarity constituents in the way that is most directly
transferable to our research question. Ideally we would routinely measure
concentrations of algal biomass, non-algal particles (including inorganic
sediment and organic sediment, I like to call this stuff dead suspended
sediment, but I guess NAP is fine), and dissolved organic carbon. These
constituents all alter light penetration in water, but they are mutually
exclusive groups with no measurement overlap between algal particles and NAP for
example. However, this is not what we usually measure. Instead, we have rich
datasets of things that could indicate our desired constituents. These map as
basically.

-   Chlorophyll *a* -\> proxy for algal biomass
-   Total suspended solids (TSS) -\> All suspended solids, subtract algal
    biomass ≅ NAP
-   DOC -\> DOC!

That means we have to basically do one key calculation. Convert chlorohpyll *a*
to biomass. Then subtract that from TSS and assume that is NAP. While this is a
absurdly simple calculation it is filled with dangers. The relationship between
chlorphyll *a* and algal biomass depends on temperature and nutrients and
species and is [not
universal](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2016GB005458).
Most studies focus on chlorohpyll:carbon ratios, but we are interested in
chlorophyll:biomass ratio, for which there are much fewer studies, though some
studies below suggest a range of 0.005 to 0.1 chlorophyll:biomass. (I need to
find the C:Biomass study I used forever ago.)

Papers that help:

-   <https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lno.10338>

-   <https://pubs.usgs.gov/twri/twri9a7/twri9a7_7.4.pdf>

-   <https://www.researchgate.net/publication/230056019_Relationship_between_Chlorophyll-a_Concentration_and_Phytoplankton_Biomass_in_Several_Reservoirs_in_Czechoslovakia/link/5cb56c164585156cd79af804/download>

-   <https://archimer.ifremer.fr/doc/2005/publication-1172.pdf>

### Approach

Our philosophy here is that we will use chlorophyll *a* directly as a proxy for
algal biomass to minimize the number of transformations to this data, since we
are primarily interested in how algae alters light attenuation, regardless of
how we estimate algae. However, we still need to calculate non-algal particle
(NAP) mass. For that we will use a range of values of chlorophyll:biomass ratio
and show the impact of picking a single ratio from a set of ranges.

*Caveats*

-   We don't have a true NAP to compare how accurate our approach is.

-   We do have Inorganic Sediment, but that is not the same thing as non-algal
    particles, since it excludes particulate organic carbon. Still, where we
    have it, we will examine the relationship between NAP and total inorganic
    sediment (TIS)

-   Not entirely sure how we propogate this uncertainty downstream? Pick an
    average? [[run them all and then get sd/error metrics to describe
    uncertainty]]

<details>

<summary>Code to apply varying algal biomass ratios.</summary>

```{r data-nap-test}

in_vis <- read_feather(file.path(data_out_dir, 'simul.feather'))

no_sech <- read_feather(file.path(data_out_dir, 'no_secchi.feather'))

range <- c(50, 100, 200, 234)


#Couldn't think of a more clever way to multiply
#chl_a by the range of values, so just made
# dataframe 4 times bigger with new column called ratio. 
nap_test <- expand_grid(no_sech, chl_ratio = range) %>%
  mutate(power = if_else(chl_ratio == 234, 0.57, 1),
         chl_a_biomass = exp(log(chl_ratio/1000)+log(chl_a)*power),
         tss_dead = tss-chl_a_biomass)
```

</details>

### Simple evaluations of conversion

#### Negative tss_dead

Probably the simplest way to see which ratios are more appropriate is to look at
how that impacts how many tss_dead observations are negative, which is
impossible.

<details>

<summary>Code for table.</summary>

```{r neg-tss-dead}

neg_tss_dead <- nap_test %>%
  mutate(negative = ifelse(tss_dead < 0, 'negative', 'positive')) %>%
  group_by(negative, chl_ratio, power) %>%
  count() %>%
  pivot_wider(names_from = 'negative',values_from = 'n')  %>%
  mutate(percent_neg = negative/(positive+negative)*100) %>%
  arrange(chl_ratio^power) %>% 
  kable(caption = "Varying chlorohpyll ratios and summary of negative tss_dead results.")

```

</details>

```{r fig-neg-tss-dead, echo=FALSE}
neg_tss_dead
```

Well, that sort of produces obvious results. Smaller ratios mean, less negative
NAP estimates, but there is a big increase between 100-200, making me think the
"correct" ratio may be in there somewhere.

#### Relationship between chlorophyll-a and tss_dead

If you recall from the previous section, there was a linear relationship between
chlorophyll *a* and TSS across all sites and water types (r^2^ \> 0.35). One
effect we'd expect to see with the tss_dead estimate is that this relationship
should decay. While there are reasons for chlorophyll *a* to be correlated with
TSS (sediment brings nutrients for example), chlorophyll *a* also includes algal
cells which are suspended particles themselves. If this approach worked well, we
should see weaker relationships between tss_dead and chlorophyll *a*.

<details>

<summary>Code for figure.</summary>

```{r fig-tss-dead-chla}

#Subset for plotting purposes
tss_dead_chla <- nap_test %>%
  #Remove negatives and very small numbers (ug/L of sediment is basically zero)
  filter(tss_dead > 0.001) %>%
  sample_frac(0.1) %>%
  mutate(chl_ratio_label = paste0("chlorohpyll a ratio: ", chl_ratio, "*chl-a^", power))  %>% 
ggplot(., aes(chl_a, tss_dead, color = type)) +
  facet_wrap(~chl_ratio_label) +
  geom_point() + 
  scale_x_log10() + 
  scale_y_log10() + 
  stat_poly_eq(small.r = T) + 
  theme_few() + 
  scale_color_manual(values = c('seagreen3','skyblue3','saddlebrown')) 

```

</details>

```{r, echo = F, fig.width = 7,fig.height = 7, fig.cap = "At low chl_ratios (10/20), there is really no change to the correlation between tss and chl_a, but at higher ratios (100/200), the relationship does breakdown some, though not completely"}

tss_dead_chla

```

-   What to do with this?
-   234\*chl^0.57^ seems logical but still high correlation in lakes between
    tss/chl_a, but that may be a real thing for lakes? (corr is actually higher
    with this approach)

### Light attenuation model exploration

Model assumptions

-   Not searching for intercept assuming observed 0.15 coefficient for intercept
    (pure water attenuation)

-   For now, not doing the middle 2.5-97.5% of data, can add back in.

-   Normality assumptions aren't fully met

<details>

<summary>Code for this section.</summary>

[[noting that kd here is 1/secchi, which is yet a different constant for
calculating kd, also for the derp in the class (me) please give insight on the
1/1.5/max secchi as kw.]]

```{r fig-kd-obs-pred, warning = F}

nap_est_no_secchi <- nap_test %>%
  ungroup() %>%
  filter(tss_dead > 0.01,
         secchi < 15 | is.na(secchi)) %>%
  mutate(secchi = ifelse(secchi < 0.1,0.1,secchi),
        kd = (1/(secchi)))

write_feather(nap_est_no_secchi %>%
                filter(chl_ratio == 234),
              file.path(data_out_dir, 'no_secchi_clean.feather'))

nap_est <- nap_est_no_secchi %>%
  filter(!is.na(secchi))


k_w <- 1/1.5/max(nap_est$secchi)

kd_mod <- function(df){
  mod <- lm((kd-k_w)  ~ 0 + tss_dead + doc + chl_a, data = df)
}

kd_resid_extract <- function(df,mod){
  df <- df %>%
    mutate(residuals = mod$residuals,
           pred = mod$fitted.values + k_w)
}

# nap_est <- nap_est %>%
#   filter(SiteID %in% fine_sites$SiteID) 

nap_mods <- group_by(nap_est, chl_ratio, power) %>%
  nest() %>%
  mutate(mods = map(data, kd_mod),
         data = map2(data, mods, kd_resid_extract))

mod234 <- nap_mods %>%
  filter(chl_ratio == 234)

nap_resid <- nap_mods %>%
  select(-mods) %>%
  unnest(data) %>%
  mutate(ratio_rmse = paste('ratio =',chl_ratio,', r^2 =', round(cor(pred,kd)^2,2)),
         year = year(date_unity))

kd_obs_pred <- nap_resid %>%
  sample_frac(0.2) %>%
  ggplot(., aes(kd, pred, color = year(date_unity))) + 
  geom_point(shape = 1) + 
  facet_wrap(~ratio_rmse) + 
  theme_few() + 
  xlab('kd (1/secchi)') + #tagging this too, as the constant of 1.
  ylab('predicted kd') +
  scale_x_log10() + 
  scale_y_log10() +
  scale_color_viridis_c() + 
  stat_smooth(method = 'lm',se = F, color = 'black') +
  geom_abline(intercept = 0, slope = 1, col = 'red') 

```

</details>

```{r fig kd-pred-obs, echo = F, fig.cap = 'Modelled light extinction coefficient (Kd) versus observed, red line is 1:1 line and blue is line of best fit, r^2 hardly varies between chlorophyll a conversion approaches.'}

kd_obs_pred

```

### Variation in slopes w/ different chl biomass ratios

-   Do all our painstaking decisions make a difference for model slopes?

-   (They didn't impact model RMSE very much)

-   No, they don't change slopes all that much either.

<details>

<summary>Code for table.</summary>

```{r pre-tab-nap-slope}

nap_slopes <- nap_mods %>%
  select(-data) %>%
  mutate(glance = map(mods, tidy)) %>%
  unnest(glance) %>%
  select(-mods,-std.error,-statistic) %>%
  pivot_wider(names_from = 'term', values_from = 'estimate') %>% 
  kable(caption = "Varying chlorophyll ratios and summary of slopes across parameters.")

# kable(nap_slopes) %>%
#   kable_paper() %>%
#   scroll_box(width = '100%', height = '600px') %>%
#   kable_styling(bootstrap_options = 'striped', full_width = T)
```

</details>

```{r tab-nap-slope, echo=F}
nap_slopes
```

[[I'm just tagging a reality check in here that I don't think there is any way
that the p value can be zero for these, additionally, each lm slope should have
it's own p value.]]

## Final Decision

-   Based on the above approaches and logic, the chl-a to biomass multiplier
    does not impact model performance very much. As such, we are going to use
    the 234\*chl_a^0.57^ log ratio going forward.
