---
output: html_document
editor_options: 
  chunk_output_type: console
  markdown:
    wrap: 80
---

# Data quality assessments

Note, also eval = F to save computation time

```{r load-pkgs-ch2, include = F}
library(sf)
library(feather)
library(GGally)
library(mapview)
library(ggpmisc)
library(Metrics)
library(kableExtra)
library(broom)
library(ggthemes)
library(knitr)
library(tidyverse)

opts_chunk$set(warning = F, message = F)
```


## Chl-a to biomass

While it is more common in coastal and oceanic research, inland water scientists
rarely measure water clarity constituents in the way that is most directly 
transferable to our research question. Ideally we would routinely measure concentrations of
algae biomass, non-algal particles
(including inorganic sediment and organic sediment, I like to call this
stuff dead suspended sediment, but I guess NAP is fine), and dissolved organic 
carbon. These constituents all alter light penetration in water, but they are 
mutually exclusive groups with no measurement overlap between algal particles and NAP
for example. However, this is not what we usually measure. Instead, we have rich 
datasets of things that could indicate our desired constituents. These map as basically.

- Chlorophyll-a -> proxy for algae biomass.
- Total suspended solids (tss) -> All suspended solids, subtract algae biomass = NAP
- DOC -> DOC!

That means we have to basically do one key calculation. Convert chl-a to biomass.
Then subtract that from tss and assume that is NAP. While this is a absurdly
simple calculation it is filled with dangers. The relationship between chl-a
and algae biomass depends on temperature and nutrients and species and is 
[not universal](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2016GB005458).
Most studies focus on Chl:C ratios, but we are interested in Chl:Biomass ratio,
for which there are much fewer studies, though some studies below suggest 
a range of 0.005 to 0.1 Chl:biomass. (I need to find the C:Biomass study I used
forever ago.)

Papers that help:

  - https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lno.10338

  - https://pubs.usgs.gov/twri/twri9a7/twri9a7_7.4.pdf

  - https://www.researchgate.net/publication/230056019_Relationship_between_Chlorophyll-a_Concentration_and_Phytoplankton_Biomass_in_Several_Reservoirs_in_Czechoslovakia/link/5cb56c164585156cd79af804/download

  - https://archimer.ifremer.fr/doc/2005/publication-1172.pdf

### Approach

Our philosphy here is that we will use chl-a directly as a proxy for algal biomass
to minimize the number of transformations to this data, since we are primarily
interested in how algae alters light attenuation, regardless of how we estimate algae.
However, we still need to calculate Non-Algal Particle mass. For that we will
use a range of values of Chl:biomass and show the impact of picking a variety 
of ranges. 

*Caveats*

 - We don't have a true NAP to compare how accurate our approach is.
 
 - We do have Inorganic Sediment, but that is not the same thing as non-algal 
 particles, since it excludes particulate organic carbon. Still, where
 we have it, we will examine the relationship between NAP and TIS (total 
 inorganic sediment). 
 
 - Not entirely sure how we propogate this uncertainty downstream? Pick an average?



```{r data-nap-test}

in_vis <- read_feather('data/out/simul.feather')

no_sech <- read_feather('data/out/no_secchi.feather')

range <- c(50, 100, 200, 234)



#Couldn't think of a more clever way to multiply
#chl_a by the range of values, so just made
# dataframe 4 times bigger with new column called ratio. 
nap_test <- expand_grid(no_sech,chl_ratio = range) %>%
  mutate(power = ifelse(chl_ratio == 234, 0.57, 1),
         chl_a_biomass = exp(log(chl_ratio/1000)+log(chl_a)*power),
         tss_dead = tss-chl_a_biomass)


```

### Simple evaluations of conversion




#### Negative tss_dead

Probably the simplest way to see which ratios are more appropriate is to
look at how that impacts how many tss_dead observations are negative (impossible).

```{r neg-tss-dead}



nap_test %>%
  mutate(negative = ifelse(tss_dead < 0, 'negative', 'positive')) %>%
  group_by(negative,chl_ratio) %>%
  count() %>%
  pivot_wider(names_from = 'negative',values_from = 'n')  %>%
  mutate(percent_neg = negative/(positive+negative)*100) %>%
  kable()

```


Well, that sort of produces obvious results. Smaller ratios mean, less negative 
NAP estimates, but there is a big increase between 100-200, making me think the
"correct" ratio may be in there somewhere. 


#### Relationship between chl-a and tss_dead

If you recall from the previous section, there was a strong linear 
relationship between chl-a and tss across all sites and water types (r2 > 0.35).
One effect we'd expect to see with the tss_dead estimate is that this relationship
should decay. While there are reasons for chl-a to be correlated with tss (sediment
brings nutrients for example), chl-a also includes algal cells which are suspended
particles themselves. If this approach worked well, we should see weaker relationships
between tss_dead and chl-a


```{r fig-tss-dead-chla, cache = T, fig.width = 7,fig.height = 7, fig.cap = "At low chl_ratios (10/20), there is really no change to the correlation between tss and chl_a, but at higher ratios (100/200), the relationship does breakdown some, though not completely"}


#Subset for plotting purposes
nap_test %>%
  #Remove negatives and very small numbers (ug/L of sediment is basically zero)
  filter(tss_dead > 0.001) %>%
  sample_frac(0.1) %>%
ggplot(., aes(chl_a,tss_dead,color = type)) +
  facet_wrap(~chl_ratio) + 
  geom_point() + 
  scale_x_log10() + 
  scale_y_log10() + 
  stat_poly_eq() + 
  theme_few() + 
  scale_color_manual(values = c('seagreen3','skyblue3','saddlebrown')) 

```


- What to do with this? 
- 234*chl^0.57 seems logical but still high correlation in lakes between tss/chl_a,
but that may be a real thing for lakes? (corr is actually higher with this approach)



### Light attenuation model exploration

Model assumptions

- Not searching for intercept assuming observed 0.15 coefficient for intercept (pure water attenuation)

- For now, not doing the middle 2.5-97.5% of data, can add back in.

- Normality assumptions aren't fully met



```{r fig-kd-obs-pred, warning = F, cache = T, fig.cap = 'modelled light extinction coefficient versus observed, red line is 1:1 line and blue is line of best fit, R2 hardly varies between chl-a conversion approaches.'}


nap_est_no_secchi <- nap_test %>%
  ungroup() %>%
  filter(tss_dead > 0.01,
         secchi < 15 | is.na(secchi)) %>%
  mutate(secchi = ifelse(secchi < 0.1,0.1,secchi),
        kd = (1/(secchi)))



write_feather(nap_est_no_secchi %>%
                filter(chl_ratio == 234),'data/out/no_secchi_clean.feather')



nap_est <- nap_est_no_secchi %>%
  filter(!is.na(secchi))




k_w <- 1/1.5/max(nap_est$secchi)


kd_mod <- function(df){
  mod <- lm((kd-k_w)  ~ 0 + tss_dead + doc + chl_a, data = df)
}


kd_resid_extract <- function(df,mod){
  df <- df %>%
    mutate(residuals = mod$residuals,
           pred = mod$fitted.values + k_w)
}

# nap_est <- nap_est %>%
#   filter(SiteID %in% fine_sites$SiteID) 


nap_mods <- group_by(nap_est,chl_ratio) %>%
  nest() %>%
  mutate(mods = map(data,kd_mod),
         data = map2(data, mods, kd_resid_extract))

mod234 <- nap_mods %>%
  filter(chl_ratio == 234)



nap_resid <- nap_mods %>%
  select(-mods) %>%
  unnest(data) %>%
  mutate(ratio_rmse = paste('ratio =',chl_ratio,', R2 =', round(cor(pred,kd)^2,2)),
         year = year(date_unity))


nap_resid %>%
  sample_frac(0.2) %>%
  ggplot(., aes(kd,pred, color = year(date_unity))) + 
  geom_point(shape = 1) + 
  facet_wrap(~ratio_rmse) + 
  theme_few() + 
  xlab('kd (1/secchi)') +
  ylab('predicted kd') +
  scale_x_log10() + 
  scale_y_log10() +
  scale_color_viridis_c() + 
  stat_smooth(method = 'lm',se = F, color = 'black') +
  geom_abline(intercept = 0, slope = 1, col = 'red') 





```





### Variation in slopes w/ different chl biomass ratios

- Do all our painstaking decisions make a difference for model slopes?

- (They didn't impact model RMSE very much)

- No, they don't change slopes all that much either. 



```{r tab-nap-slope}

nap_slopes <- nap_mods %>%
  select(-data) %>%
  mutate(glance = map(mods, tidy)) %>%
  unnest(glance) %>%
  select(-mods,-std.error,-statistic) %>%
  pivot_wider(names_from = 'term', values_from = 'estimate')



kable(nap_slopes) %>%
  kable_paper() %>%
  scroll_box(width = '100%', height = '600px') %>%
  kable_styling(bootstrap_options = 'striped', full_width = )
```


# Final Decision

- Based on the above approaches and logic, the chl-a to biomass multiplier does 
not impact model performance very much. As such, we are going to use the 234 log 
ratio going forward. 



