---
output: html_document
editor_options: 
  chunk_output_type: console
---



# Data quality assessments

Note, also eval = F to save computation time

```{r, include = F}

library(tidyverse)
library(sf)
library(feather)
library(GGally)
library(mapview)
library(ggpmisc)
library(lubridate)
library(Metrics)
library(kableExtra)
library(broom)
library(ggthemes)
library(ggtern)
library(tmap)

knitr::opts_chunk$set(warning = F, message = F)
```

# Data read-and-munge



```{r}


simul <- read_feather('data/out/simul.feather') %>%
  #rename(chl_a = chla) %>%
  filter(if_all(c(chl_a,doc,secchi,tss), ~ . >= 0 )) %>%
  dplyr::filter(if_all(c(chl_a,doc,secchi,tss), 
                       ~!is.na(.) & . > quantile(.,0.001) & 
                         . < quantile(.,0.999))) %>%
  mutate(type = ifelse(grepl('Lake',type),'Lake',type))


range <- c(234)


max_secchi <- max(simul$secchi)
#Couldn't think of a more clever way to multiply
#chl_a by the range of values, so just made
# dataframe 4 times bigger with new column called ratio. 
nap_234 <- expand_grid(simul,chl_ratio = range) %>%
  mutate(power = ifelse(chl_ratio == 234, 0.57, 1),
         chl_a_biomass = exp(log(chl_ratio/1000)+log(chl_a)*power),
         tss_dead = tss-chl_a_biomass,
         chl_a = chl_a) %>%
  filter(chl_ratio == 234) %>%
  mutate(secchi = round(ifelse(secchi < .01,.01,secchi),2),
         kd = round((1/(secchi)),5))



no_sdd <- read_feather('data/out/no_secchi.feather') %>%
  #rename(chl_a = chla) %>%
  filter(chl_a <= max(simul$chl_a, na.rm  = T) &
           chl_a >= min(simul$chl_a, na.rm = T),
         tss <= max(simul$tss, na.rm  = T) &
           tss >= min(simul$tss, na.rm = T),
         doc <= max(simul$doc, na.rm  = T) &
           doc >= min(simul$doc, na.rm = T),
         secchi <= max(simul$secchi,na.rm =T) &
           secchi >= min(simul$secchi, na.rm = T) |
           is.na(secchi))%>%
  mutate(type = ifelse(grepl('Lake',type),'Lake',type)) %>%
  expand_grid(.,chl_ratio = range) %>%
  mutate(power = ifelse(chl_ratio == 234, 0.57, 1),
         chl_a_biomass = exp(log(chl_ratio/1000)+log(chl_a)*power),
         tss_dead = tss-chl_a_biomass) %>%
  filter(chl_ratio == 234) %>%
  filter(tss_dead > 0.01)  
```


# Modelling development

## Model Setup

```{r}
k_w <- 1/1.5/max_secchi


kd_mod <- function(df){
  mod <- lm((kd-k_w)  ~ 0 + (tss_dead + doc + chl_a), data = df)
}


kd_resid_extract <- function(df,mod){
  df <- df %>%
    mutate(residuals = mod$residuals,
           pred = mod$fitted.values + k_w)
}

tidier <- function(mod){
  
  kd <- mod$model$`(kd - k_w)`+k_w
  pred <- mod$fitted.values + k_w
  
  r2_real = cor(kd,pred)^2
  tidied_mod <- tidy(mod) %>%
    mutate(r2 = r2_real)
}



```

## Modelling random subsets of data


```{r}


# Ten random iterations, a bit overkill for sure. 
group_sizes <- rep(round(nrow(nap_234)/c(50,100,250,500,1000,2000,5000,10000, nrow(nap_234)),0),10)

random_coef_holder <- list()

## abandoned parallel version. 

# plan(multisession, workers = 5) 
# future_map(.x = group_sizes,
#            .options = furrr_options(seed = 123),
#            .f = ~ {
#              nap_234$rgroup <- sample(1:.x,
#                                       size = nrow(nap_234),
#                                       replace = T)
#              
#                nap_234 %>%
#                  mutate(group_size = .x) %>%
#                  group_by(chl_ratio, rgroup) %>%
#                  nest() %>%
#                  mutate(
#                    mods = map(data, kd_mod),
#                    data = map2(data, mods, kd_resid_extract),
#                    tidy_mod = map(mods, tidier)
#                  ) %>%
#                  select(-mods, -data) %>%
#                  unnest(tidy_mod) %>%
#                  mutate(group_size = .x,
#                         each_group = round(nrow(nap_234) / group_size),0)
#            })

for(i in 1:length(group_sizes)){
  nap_234$rgroup <- sample(1:group_sizes[i], 
                           size = nrow(nap_234), 
                           replace = T)
  
  random_coef_holder[[i]] <- nap_234 %>%
    mutate(group_size = group_sizes[i]) %>%
    group_by(chl_ratio, rgroup) %>%
    nest() %>%
    mutate(mods = map(data,kd_mod),
           data = map2(data, mods, kd_resid_extract),
           tidy_mod = map(mods, tidier)) %>%
    select(-mods, -data) %>%
    unnest(tidy_mod) %>%
    mutate(group_size = group_sizes[i],
           each_group = round(nrow(nap_234)/group_size),0)
}

all_random_mods <- do.call('rbind',random_coef_holder) 


cv_estimate <- all_random_mods %>%
  group_by(each_group,term, group_size) %>%
  summarize(cv = sd(estimate)/mean(estimate)) %>%
  pivot_wider(names_from = term,
              values_from = cv)



ggplot(all_random_mods,
       aes(y = estimate, x = as.factor(each_group))) + 
  geom_boxplot() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) 



```






## Modelling over different ecoregions and water types

- Ecoregion Level 2
- Type

### Ecoregion prep

```{r, eval = F}
download.file('https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/cec_na/na_cec_eco_l2.zip',
              destfile = 'data/in/eco2.zip')

download.file('https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l3.zip',
              destfile = 'data/in/eco3.zip')

download.file('https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l4.zip',
              destfile = 'data/in/eco4.zip')

unzip('data/in/eco1.zip',exdir = 'data/in/eco1')

unzip('data/in/eco2.zip',exdir = 'data/in/eco2')

unzip('data/in/eco3.zip',exdir = 'data/in/eco3')

unzip('data/in/eco4.zip', exdir = 'data/in/eco4')


```


### Eco Joinery

```{r}
eco2 <- st_read('data/in/eco2/NA_CEC_Eco_Level2.shp') %>%
  select(code = NA_L2CODE,
         name = NA_L2NAME) %>%
  filter(name != 'WATER')


eco3 <- st_read('data/in/eco3/us_eco_l3.shp') %>%
  select(l3code = NA_L3CODE,
         l3name = NA_L3NAME,
         code = NA_L2CODE,
         name = NA_L2NAME) %>%
  filter(name != 'WATER')

eco4 <- st_read('data/in/eco4/us_eco_l4_no_st.shp')  %>%
  select(l3code = NA_L3CODE,
         l3name = NA_L3NAME,
         code = NA_L2CODE,
         name = NA_L2NAME,
         l4code = US_L4CODE,
         l4name = US_L4NAME)




eco_joinery <- function(df = nap_234,
                        eco_reg = eco2,
                        code = code){
  
  
  nap_site_pass <- df %>%
    distinct(SiteID, lat, long) %>%
    st_as_sf(., coords = c('long','lat'), crs = 4326) %>%
    st_transform((st_crs(eco_reg))) %>%
    st_join(eco_reg)
  
  nap_nas <- nap_site_pass %>%
    filter(is.na({{code}})) %>%
    select(SiteID) %>%
    st_join(eco_reg, join = st_nearest_feature) 
  
  nap_sites <- nap_site_pass %>%
    filter(!is.na({{code}})) %>%
    rbind(nap_nas)
  
  
  return_df <- inner_join(nap_sites,df) %>%
    st_drop_geometry()
  
}


zone_min <- 150


nap_234_eco4_all <- eco_joinery(df = nap_234,
                            eco_reg = eco4,
                            code = l4code) %>%
  group_by(l4name,l4code) %>%
  add_count(name = 'l4count') 

no_sdd_4 <- eco_joinery(df = no_sdd,
                            eco_reg = eco4,
                            code = l4code) %>%
  group_by(l4name,l4code) %>%
  add_count(name = 'l4count') 


no_sdd_4_3 <-  no_sdd_4 %>%
  mutate(zone = 3) %>%
  group_by(l3code,l3name) %>%
  add_count(name = 'l3count') %>%
  ungroup() %>%
  filter(l3count > zone_min) %>%
  select(names(no_sdd), code = l3code, name = l3name) 




nap_234_eco4_eco3 <- nap_234_eco4_all %>%
  mutate(zone = 3) %>%
  group_by(l3code,l3name) %>%
  add_count(name = 'l3count') %>%
  ungroup() %>%
  filter(l3count > zone_min) %>%
  select(names(nap_234), code = l3code, name = l3name) 

nap_234_eco4 <- nap_234_eco4_all %>%
    filter(l4count > zone_min) %>%
  mutate(zone = 4) %>%
  select(names(nap_234),code = l4code, name = l4name) 




no_4zone <- nap_234 %>%
  anti_join(nap_234_eco4)


nap_234_eco3 <- eco_joinery(df = no_4zone,
                            eco_reg = eco3,
                            code = l3code) %>%
  group_by(l3code,l3name) %>%
  add_count(name = 'l3count') %>%
  filter(l3count > zone_min) %>%
  mutate(zone = 3) %>%
  select(names(nap_234),code = l3code, name = l3name)


no_3zone <- no_4zone %>%
  anti_join(nap_234_eco3)


nap_234_eco2 <- eco_joinery(df = no_3zone,
                            eco_reg = eco2,
                            code = code) %>%
  mutate(zone = 2) %>%
  select(names(nap_234),code = code, name = name)



zoned_4 <- bind_rows(nap_234_eco2,nap_234_eco3,nap_234_eco4) %>%
  group_by(name,code) %>%
  add_count() %>%
  filter(n > zone_min)

zoned_3 <- bind_rows(nap_234_eco2, nap_234_eco3, nap_234_eco4_eco3) %>%
  group_by(name,code) %>%
  add_count() %>%
  filter(n > zone_min)


zoned_2 <-  eco_joinery(df = nap_234,
                            eco_reg = eco2,
                            code = code) %>%
  group_by(name,code) %>%
  add_count() %>%
  filter(n > zone_min) %>%
  select(names(nap_234), code = code, name = name) 

```



### Maps


```{r, eval = T}

simple_3 <- eco3

zoned_3_site_counts <- zoned_3 %>%
  group_by(code,name) %>%
  count()



simple_3_3_join<- simple_3 %>% 
  filter(l3code %in% zoned_3_site_counts$code) %>%
  group_by(code = l3code,name = l3name) %>%
  summarize() %>%
  inner_join(zoned_3_site_counts)

## DO ABOVE WITH Z2

simple_3_2_join<- simple_3 %>% 
  filter(code %in% zoned_3_site_counts$code) %>%
  group_by(code,name) %>%
  summarize() %>%
  inner_join(zoned_3_site_counts) %>%
  st_difference(.,
                st_union(simple_3_3_join))



modelable_z3 <- bind_rows(simple_3_3_join,
                          simple_3_2_join) 


mapview(modelable_z3, zcol = 'n')

no_sdd_modelable <- no_sdd %>%
  distinct(SiteID, lat, long) %>%
  st_as_sf(., coords = c('long','lat'), crs = 4326) %>%
  st_transform((st_crs(modelable_z3))) %>%
  st_join(modelable_z3)




no_sdd_na <- no_sdd_modelable %>%
    filter(is.na(code)) %>%
    select(SiteID) 


# SLOW!
dist_50 <- nngeo::st_nn(no_sdd_na, modelable_z3, k = 1, returnDist = TRUE)


no_sdd_100 <- no_sdd_na %>%
  mutate(mindex = unlist(dist_50$nn),
                dist = unlist(dist_50$dist)) %>%
  filter(dist <= 100)


no_sdd_final <- bind_rows(no_sdd_modelable %>%
                            filter(!is.na(code)),
                              st_join(no_sdd_100,modelable_z3, 
                                      st_nearest_feature)) %>%
  filter(!is.na(code)) %>%
  group_by(SiteID) %>%
  add_count(name = 'codenames') 




no_sdd_z3 <- inner_join(no_sdd, no_sdd_final)


```




# Model Application

## Global Model

```{r}
## Global model




nap_mods <- group_by(nap_234,chl_ratio) %>%
  nest() %>%
  mutate(mods = map(data,kd_mod),
         data = map2(data, mods, kd_resid_extract))

mod234 <- nap_mods %>%
  dplyr::filter(chl_ratio == 234)

mod <- mod234$mods[[1]]


with_sech <- nap_234 %>%
  mutate(kd_pred = predict(mod,.),
         kd_pred = kd_pred + k_w)

r2 = cor(with_sech$kd,with_sech$kd_pred)^2



sech_val <- ggplot(with_sech,aes(x = kd, y = kd_pred, color = type)) + 
  geom_point(shape = 1) + 
  scale_x_log10() + 
  scale_y_log10() + 
  theme_few() +
  scale_color_few() + 
  xlab('Actual kd from Secchi') + 
  ylab('Predicted kd') +
  ggtitle('Modeled Light Extinction R2 = 0.34') + 
  geom_abline(slope = 1) 



ggsave(plot = sech_val,
       filename = 'figs_2023/sech_val_old.png',
       dpi=300,
       height = 6,
       width = 6)

```



## Granular models

- No P value filtering
- Applying mods with logic that ecoregion can modify both intercept and slope
- Applying single models (e.g. lm(kd~(tss+secchi+chla)*type))

### Type mods


```{r}
#Single type mod

type_mod <- lm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):type, data = nap_234)


pred_mutation <- function(df = nap_234, smod = type_mod) {
  df %>%
  mutate(kd_pred = predict(smod, .) + k_w)
         # tss_kd = ((smod$coefficients[1])*tss_dead)/kd_pred,
         # doc_kd = ((smod$coefficients[2])*doc)/kd_pred,
         # chl_kd = ((smod$coefficients[3])*chl_a)/kd_pred,
         # test_1 = tss_kd + doc_kd + chl_kd) 
}

tidy_type <- tidier(type_mod)

tidy_type
```

#### Nested Type mods (old)

```{r}
# nested_type_mods
nap_type_mods <- nap_234 %>%
  group_by(type) %>%
  add_count() %>%
  group_by(type,n) %>%
  nest() %>%
  mutate(mods = map(data,kd_mod),
         tidy_mod = map(mods, tidier))

nap_type_tidy <- nap_type_mods %>%
  select(-mods,-data)  %>%
  unnest(tidy_mod)
```


### Eco Mods


### Single Eco 3 mod

- 

```{r}

eco2_type_glm <- glm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):code:type, data = zoned_2) 

eco2_glm <- glm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):code, data = zoned_2) 


eco3_type_glm <- glm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):code:type, data = zoned_2) 


eco3_type_glm <- glm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):code:type, data = zoned_3) 

eco3_glm <- glm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):code, data = zoned_3 %>%
                  filter(n > zone_min)) 

eco4_glm <- glm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):code, data = zoned_4) 

eco4_type_glm <- glm((kd-k_w) ~ 0 + (tss_dead + doc + chl_a):code:type,
                data = zoned_4)

BIC(eco2_glm)
BIC(eco2_type_glm)
BIC(eco3_glm)
BIC(eco4_glm)
BIC(eco3_type_glm)
BIC(eco4_type_glm)

summary_mods <- list(eco2_glm, eco2_type_glm, eco3_glm, eco3_type_glm,
             eco4_glm, eco4_type_glm)


## TO DO!

## Look at spatial of variation in estimates

## Apply estimates to NO_SDD Data

## Clean up modelling and spatial joining pipeline. 
tidy_eco3 <- tidier(eco3_glm) 




ggplot(tidy_eco3 %>%
         filter(grepl('doc',term)), aes(x = estimate) ) +
  geom_histogram() 


fit_comparer = function(mod = eco2_glm){
  
  kd <- mod$model$`(kd - k_w)`+k_w
  pred <- mod$fitted.values + k_w
  
  r2_real = cor(kd,pred)^2
  bic = BIC(mod)
  aic = AIC(mod)
  
  tibble(r2 = r2_real,
         bic = bic,
         aic = aic)
}

tidier_mods <- map_dfr(summary_mods, fit_comparer) 

tidy_4_type <- tidier(eco3_glm) %>%
  filter(!is.na(estimate))

tidier_mods

```


# Model application zone

```{r}
#applying mods to no_sdd_z3. 
?str_split_1

#Step one split out model coefs by eco region and parameter

coef_splitter <- function(df = tidy_eco3) {
  df %>%
    separate_wider_delim(term, delim = ":code", names = c("parameter", "code")) %>%
    select(code,parameter,estimate) %>%
    pivot_wider(id_cols = code,names_from = parameter,
                names_prefix = 'est_',
                values_from = estimate)
}


final_coefs <- no_sdd_z3 %>%
  mutate(kd = round((1/(secchi)),5),
        kd_pred = predict(eco3_glm,.) + k_w) %>%
  inner_join(., coef_splitter(tidy_eco3)) %>%
  mutate(tss_kd = (est_tss_dead*tss_dead)/kd_pred,
         doc_kd = (est_doc*doc)/kd_pred,
         chl_kd = (est_chl_a*chl_a)/kd_pred,
         test_1 = tss_kd + doc_kd + chl_kd) %>%
  select(-n,-mindex, -dist) %>%
  mutate(month = month(date_unity),
         year = year(date_unity))



```


### Various summaries


```{r}



final_monthly_means <- final_coefs %>% 
  group_by(SiteID, lat, long, type, code, name, month) %>%
  select(tss_dead, tss, doc, chl_a,kd, kd_pred, tss_kd, doc_kd, chl_kd) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm=T)),
            n = n())
  

final_yearly_means <- final_coefs %>%
  group_by(SiteID, lat, long, type, code, name, year) %>%
  add_count() %>%
  select(tss_dead, tss, doc, chl_a,kd, kd_pred, tss_kd, doc_kd, chl_kd) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm=T)),
            n = n())

write_csv(final_coefs, file = 'data/final/final_coefs.csv')
write_csv(final_monthly_means, file = 'data/final/final_monthly.csv')
write_csv(final_yearly_means, file = 'data/final/final_yearly.csv')

```






### Tern Test

```{r}
library(ggtern)


ggtern(final_coefs, aes(chl_kd,doc_kd,tss_kd,color = type)) + 
  geom_point(shape = 1, size = 0.5) +
  scale_color_manual(values = c('tan3','blue4','gray60')) + 
  theme_bw() + 
  theme_showarrows()  +
  labs(x = 'Algae', 
       y = 'DOC',
       z = 'TSS')  

```





### Performance 

```{r}

ggplot(final_coefs,aes(x = kd, y = kd_pred)) + 
  geom_point() + 
  scale_y_log10() + 
  scale_x_log10()

```


### Coefficient extraction

```{r}
df = tidy_eco3



```








```{r}
nap_eco_mods <- zoned %>%
  group_by(name,code,n) %>%
  nest() %>%
  mutate(mods = map(data,kd_mod),
         tidy_mod = map(mods, tidier))  %>%
  select(-data)

nap_eco_coefs <- nap_eco_mods %>%
  select(-mods) %>%
  unnest(tidy_mod) %>%
  mutate(eco2 = abbreviate(name,4),
         N = log10(n)) %>%
  filter(n >= 20) %>%
  mutate(label = paste0(eco2,' (',n,')'),
         label = fct_reorder(label,n))


global_mod <- tidier(mod)




eco_estimates <- ggplot(nap_eco_coefs,aes(x = fct_reorder(label, n), y = estimate, size = N)) + 
  geom_hline(data = nap_type_tidy,aes(yintercept = estimate),
             linetype = 2) + 
  geom_hline(data = global_mod, aes(yintercept = estimate), color = 'black') + 
  geom_point(shape = 1) + 
  facet_wrap(~term,ncol = 1) + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
  scale_size(range = c(0.5,5), breaks = c(1,2,3,4),
             labels = c('10','100','1000','10,000')) + 
  theme_few() + 
  scale_color_few() + 
  xlab('Ecoregion Level 2 or 3')


eco_estimates

ggsave(plot = eco_estimates,
       filename = 'figs_2023/eco_type_estimates_fixed.png',
       dpi=300,
       height = 5,
       width = 7)


## Add envelope and organize by counts. 
```


# Methods Figure 1

Envelope of random versus eco region

## Cones


```{r}
random_99s <- all_random_mods %>%
  group_by(group_size, term, each_group) %>%
  summarize(q99 = quantile(estimate,.99999),
            q01 = quantile(estimate, .00001),
            max99 = max(estimate),
            min99 = min(estimate))

random_99s

ggplot() +
  geom_ribbon(data = random_99s, aes(x = each_group, ymin = min99, ymax=max99),
              color = 'gray20') + 
  geom_point(data = nap_eco_coefs,
             aes(x = n, y = estimate, color = label)) + 
  facet_wrap(~term, nrow = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  scale_x_log10(labels = scales::label_log(digits = 1)) + 
  theme_few() + 
  xlab('Group size') + 
  ylab('kd estimate')



```

## Estimate Maps

```{r}

simple_4 <- st_simplify(eco4, dTolerance = 1000) %>%
  st_transform(4326) 



names(nap_eco_coefs)
names(simple_4$l4code)


s4 <- inner_join(simple_4,nap_eco_coefs, by = join_by(l4code == code), relationship = 'many-to-many')

s3 <- inner_join(simple_4,nap_eco_coefs, by = join_by(l3code == code), relationship = 'many-to-many')

s2 <- inner_join(simple_4,nap_eco_coefs, by = join_by(code == code), relationship = 'many-to-many')




regions <- bind_rows(s2,s3) %>%
  bind_rows(s4)

regions %>%
  filter(term == 'doc') %>%
  mapview(., zcol = 'estimate')


```




#### Methods figure

```{r}


sd_mean <- function(x){
  sd(x)/mean(x)
}

means <- nap_234_eco %>%
  mutate(abrev = abbreviate(name, 5)) %>%
  group_by(type, name, code,abrev) %>%
  summarize((across(where(is.numeric), sd_mean))) %>%
  select(name,abrev,chl_a,tss_dead,doc)




```


#### R2 


```{r}

nap_organized <- nap_eco_coefs %>%
  select(code, name, n, term, estimate, r2) %>% 
  right_join(global_mod %>%
               select(global_r2 = r2,
                      global_est = estimate,
                      term),
             by = c('term')) %>%
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
  mutate(portion = n/nrow(zoned)) %>%
  arrange(-r2) 
  



nap_organized %>%
kable(.) %>%
  kable_styling(.)



```




### Apply to no sech data. 

```{r, include = F}




d234 <- no_sdd %>%
  mutate(kd_pred = predict(mod, .) + k_w,
         tss_kd = ((mod$coefficients[1])*tss_dead)/kd_pred,
         doc_kd = ((mod$coefficients[2])*doc)/kd_pred,
         chl_kd = ((mod$coefficients[3])*chl_a)/kd_pred,
         test_1 = tss_kd + doc_kd + chl_kd) 


```

# Results



```{r}

no_sech_tern_global_mod <- ggtern(d234 , 
                                  aes(chl_kd,doc_kd,tss_kd,color = type)) + 
  geom_point(shape = 1, size = 0.5) +
  scale_color_manual(values = c('tan3','blue4','gray60')) + 
  theme_bw() + 
  theme_showarrows()  +
  labs(x = 'Algae', 
       y = 'DOC',
       z = 'TSS')  

no_sech_tern_global_mod


```

## Granula mod tern


```{r}

coanalyst <- function(df, local_mod){
  df = df %>%
    mutate(kd_pred = predict(local_mod, df) + k_w,
           tss_kd = ((local_mod$coefficients[1])*tss_dead)/kd_pred,
           doc_kd = ((local_mod$coefficients[2])*doc)/kd_pred,
           chl_kd = ((local_mod$coefficients[3])*chl_a)/kd_pred,
           test_1 = tss_kd + doc_kd + chl_kd) 
}






gran_mods <- no_sdd_eco %>%
  group_by(code, name, type) %>%
  nest() %>%
  inner_join(nap_eco_mods) %>%
  filter( n > 10) %>%
  mutate(type_eco_preds = map2(data, mods, coanalyst)) %>%
  select(code, name, type, type_eco_preds) %>%
  unnest(type_eco_preds) %>%
  arrange(date_unity)


granula_tern <- ggtern(gran_mods , 
                       aes(chl_kd,doc_kd,tss_kd,color = type)) + 
  geom_point(shape = 1, size = 0.5) +
  scale_color_manual(values = c('tan3','blue4','gray60')) + 
  theme_bw() + 
  theme_showarrows()  +
  labs(x = 'Algae', 
       y = 'DOC',
       z = 'TSS')  

granula_tern


ggsave(plot = granula_tern,
       filename = 'figs_2023/granular_models.png',
       dpi=300,
       height = 6,
       width = 6)


summary(gran_mods)


gran_err <- gran_mods %>%
  filter(!is.na(secchi)) %>%
  mutate(kd = round((1/(secchi)),5)) %>%
  filter(tss_kd > 0,
         doc_kd > 0,
         chl_kd > 0) %>%
  filter(secchi <= max(nap_234$secchi),
         secchi >= min(nap_234$secchi))

r2_gran = cor(gran_err$kd, gran_err$kd_pred)^2

gran_rmse = rmse(gran_err$kd, gran_err$kd_pred)

gran_mape = mape(gran_err$kd, gran_err$kd_pred)

```



### Ternaries by type with Jim's Hex zones

```{r}


# 
type_tern_sums <- d234 %>%
  group_by(type) %>%
  select(tss_kd,doc_kd,chl_kd,type) %>%
  summarize(across(where(is.numeric), mean))




tern_elip <- ggtern(d234,
                    aes(chl_kd,doc_kd,tss_kd)) + 
  geom_point(shape = 1, size = 0.1,
             color = 'gray30') +
  scale_color_manual(values = c('tan3','blue4','gray60')) + 
  theme_bw() + 
  theme_showarrows()  +
  labs(x = 'Algae', 
       y = 'DOC',
       z = 'TSS')  + 
  facet_wrap(~type,
             ncol = 3) + 
  geom_mean_ellipse() + 
  #geom_hex_tern(bins = 50) + 
  geom_point(data = type_tern_sums, color = 'white') 


tern_elip

ggsave(plot = tern_elip,
       filename = 'figs_2023/tern_elip_new.png',
       dpi=300,
       height = 9,
       width = 4)

```




To have results we need to find some interesting candidate sites to explore. These
will be places that either 1) have lots of repeat data (temporal variation) or 2) 
have lots of spatial data (spatial variation), or ideally both!


## Temporal Variation


### Sites with long-term records

Long-term means:

- At least 10 years of data with at least 3 observations in each year


```{r}

library(ggtern)


long_terms <- d234 %>%
  group_by(lat,long,SiteID) %>%
  distinct(date,.keep_all = T) %>%
  mutate(year = year(date)) %>%
  mutate(n_years = n_distinct(year)) %>%
  filter(n_years > 5) %>%
  mutate(range = paste(min(year),max(year)))

# ggtern(d234, aes(tss_kd,doc_kd,chl_kd,
#                  size = secchi_pred)) + 
#     geom_point(shape = 19) +
#     scale_color_viridis_c() + 
#     theme_bw() + 
#     scale_size(range = c(0.1,3)) + 
#     theme_showarrows() + 
#     labs(x = 'TSS', 
#          y = 'DOC',
#          z = 'Algae') 

year_plotter <- function(df = nested_long$data[[1]]){
  ggtern(df, aes(tss_kd,doc_kd,chl_kd,color = year,
                 size = secchi_pred)) + 
    geom_point(shape = 19) +
    scale_color_viridis_c() + 
    theme_bw() + 
    scale_size(range = c(0.5,2)) + 
    theme_showarrows() + 
    labs(x = 'TSS', 
         y = 'DOC',
         z = 'Algae') + 
    ggtitle(paste(unique(df$id)))
}

nested_long <- long_terms %>%
  group_by(lat,long,SiteID, year, type) %>%
  mutate(n = n()) %>%
  filter(n > 2) %>%
  summarize(across(where(is.numeric), mean)) %>%
  group_by(lat, long, SiteID, type) %>%
  mutate(id = SiteID) %>%
  nest() %>%
  ungroup() %>%
  mutate(mean_plots = map(data, year_plotter))


site_sf <- long_terms %>%
  distinct(lat,long,SiteID,range, n_years) %>% 
  inner_join(nested_long) %>%
  st_as_sf(.,coords = c('lon','lat'), crs = 4326) %>%
  ungroup() 


# mapviewOptions(fgb = F)
# m1 <- mapview(site_sf,
#               zcol = 'n_years',
#               popup = leafpop::popupGraph(site_sf$mean_plots))
# 
# 
# 
# 
# mapshot(m1, url = 'long_terms.html')
```



### Seasonal variation


```{r}


season_sites <- d234 %>%
  filter(SiteID %in% site_sf$SiteID) %>%
  mutate(year = year(date_unity),
         month = month(date_unity)) %>%
  group_by(lat,long,SiteID,type,year) %>%
  summarize(dist_months = n_distinct(month)) %>%
  filter(dist_months > 5)




month_nest <- d234 %>%
  filter(SiteID %in% season_sites$SiteID) %>%
  mutate(year = year(date_unity),
         month = month(date_unity)) %>%
  group_by(SiteID, lat, long, month) %>%
  summarize(across(where(is.numeric),mean)) %>%
  group_by(SiteID, lat, lon) %>%
  mutate(id = SiteID) %>%
  nest()



month_plotter <- function(df =  month_nest$data[[1]]){
  ggtern(df, aes(tss_kd,doc_kd,chl_kd,color = month,
                 size = secchi_pred)) + 
    theme_bw() + 
    theme_showarrows() + 
    labs(x = 'TSS', 
         y = 'DOC',
         z = 'Alg') + 
    geom_point() + 
    scale_color_gradient2(low = 'green3',
                          mid = 'gray20',
                          high = 'purple3',
                          midpoint = 6) + 
    scale_size(range = c(0.5,3))
}

month_sf <- month_nest %>%
  st_as_sf(.,coords = c('lon','lat'), crs = 4326) %>%
  ungroup()  %>%
  mutate(month_plots = map(data, month_plotter)) 

# mapviewOptions(fgb = F)
# month_map <- mapview(month_sf,
#                      popup = leafpop::popupGraph(month_sf$month_plots))
# 
# mapshot(month_map, url = 'month_lon_terms.html')

```


## Type means


```{r}
month_site_means <- d234 %>%
  mutate(month = month(date_unity),
         year = year(date_unity),
         season = cut(month,4)) %>%
  group_by(lat, long, SiteID, season, type) %>%
  summarize(across(where(is.numeric), mean))


ggtern(month_site_means, aes(tss_kd,doc_kd,chl_kd,
                             size = secchi_pred)) + 
  theme_bw() + 
  theme_showarrows() + 
  labs(x = 'TSS', 
       y = 'DOC',
       z = 'Alg') + 
  geom_point() + 
  scale_color_gradient2(low = 'green3',
                        mid = 'gray20',
                        high = 'purple3',
                        midpoint = 6) + 
  scale_size(range = c(0.5,3)) + 
  facet_grid(season ~ type)



```



# MC/JJ Questions

- Exemplars for seasonal/long-term/temporal to show off ideas 

- Here's everything mapped over where we have tss/chl_a/doc and clustered?

- When making results drop secchi requirement and model secchi. 

- Huc membership

- Screening for long-term data at least 1/2 per year for > 20 years

- intra-annual at least 5 months per year. 

## Spatial Variation


```{r}


fox_ids <- tibble(SiteID = c('USGS-04084422','USGS-04084445','USGS-04084475','USGS-04085054',
                             'USGS-04085059', 'USGS-04085139'),
                  order = c(6,5,4,3,2,1))



fox_r <- d234  %>%
  inner_join(fox_ids) %>%
  mutate(date = as.Date(date_unity)) 


fox_r_map <- fox_r %>%
  distinct(SiteID, .keep_all = T) %>%
  st_as_sf(.,coords = c('lon','lat'),crs = 4326)

mapview(fox_r_map)

summary(fox_r)
cool_dates <- fox_r %>%
  group_by(date) %>%
  count() %>%
  arrange(-n) %>%
  ungroup() %>%
  slice(1:5) %>%
  arrange(date)



fox_days <- fox_r %>%
  dplyr::filter(date %in% cool_dates$date) 




ggtern(fox_days,aes(tss_kd,chl_kd,doc_kd, color = order)) + 
  geom_point() + 
  facet_wrap(~date,ncol=3) + 
  theme_bw() + 
  theme_showarrows() + 
  labs(x = 'TSS', 
       y = 'DOC',
       z = 'Alg') + 
  geom_point() + 
  geom_line() +
  scale_color_gradient(low = 'skyblue',high = 'blue3') + 
  theme(legend.position = c(0.8,0.2))


```


### Additions

- Geographic Coverage

- Plausible space (not deep clarity and high chl-a)

- Switch model application to wherever we have all data.

- Requires a reasonable R2 to predict KD. Highly skewed in 
the data

- R2 round the corner. 



