---
output: html_document
editor_options: 
  chunk_output_type: console
---



# Data quality assessments

Note, also eval = F to save computation time

```{r, include = F}

library(tidyverse)
library(sf)
library(feather)
library(GGally)
library(mapview)
library(ggpmisc)
library(lubridate)
library(Metrics)
library(kableExtra)
library(broom)
library(ggthemes)
library(ggtern)
library(devtools)
library(glue)



knitr::opts_chunk$set(warning = F, message = F)
```

## New stricter data

```{r}



simul <- read_feather('data/out/simul.feather') %>%
  #rename(chl_a = chla) %>%
  filter(if_all(c(chl_a,doc,secchi,tss), ~ . > 0 )) %>%
  dplyr::filter(if_all(c(chl_a,doc,secchi,tss), 
                       ~!is.na(.) & . > quantile(.,0.001) & 
                         . < quantile(.,0.999))) %>%
  mutate(type = ifelse(grepl('Lake',type),'Lake',type))



range <- c(234)

summary(simul$chl_a)
max_secchi <- max(simul$secchi)
#Couldn't think of a more clever way to multiply
#chl_a by the range of values, so just made
# dataframe 4 times bigger with new column called ratio. 
nap_234 <- expand_grid(simul,chl_ratio = range) %>%
  mutate(power = ifelse(chl_ratio == 234, 0.57, 1),
         chl_a_biomass = exp(log(chl_ratio/1000)+log(chl_a)*power),
         tss_dead = tss-chl_a_biomass) %>%
  filter(chl_ratio == 234) %>%
  filter(tss_dead > 0.01,
         secchi < max_secchi | is.na(secchi)) %>%
  mutate(secchi = round(ifelse(secchi < .01,.01,secchi),2),
        kd = round((1/(secchi)),5))
```


## Model Application

```{r}
k_w <- 1/1.5/max_secchi


kd_mod <- function(df){
  mod <- lm((kd-k_w)  ~ 0 + tss_dead + doc + chl_a, data = df)
}


kd_resid_extract <- function(df,mod){
  df <- df %>%
    mutate(residuals = mod$residuals,
           pred = mod$fitted.values + k_w)
}



nap_mods <- group_by(nap_234,chl_ratio) %>%
  nest() %>%
  mutate(mods = map(data,kd_mod),
         data = map2(data, mods, kd_resid_extract))

mod234 <- nap_mods %>%
  dplyr::filter(chl_ratio == 234)

mod <- mod234 %>%
  pull(mods) %>%
  .[[1]]

k_w <- 1/1.5/max_secchi




with_sech <- nap_234 %>%
  mutate(kd_pred = predict(mod,.),
         kd_pred = kd_pred + k_w)

r2 = round(cor(with_sech$kd,with_sech$kd_pred)^2, 2)

sech_val <- ggplot(with_sech,aes(x = kd, y = kd_pred, color = type)) + 
  geom_point(shape = 1) + 
  scale_x_log10() + 
  scale_y_log10() + 
  theme_few() +
  scale_color_few() + 
  xlab('Actual kd from Secchi') + 
  ylab('Predicted kd') +
  ggtitle(glue('Modeled K, $R^{2}$ = {r2}')) + 
  geom_abline(slope = 1) 


sech_val


ggsave(plot = sech_val,
       filename = 'figs_2023/sech_val.png',
       dpi=300,
       height = 6,
       width = 6)

```




# Model performance
```{r}

rmse(with_sech$kd,with_sech$kd_pred)
mape(with_sech$kd,with_sech$kd_pred)



```


```{r, include = F}


no_sdd <- read_feather('data/out/no_secchi.feather') %>%
  #rename(chl_a = chla) %>%
  filter(chl_a <= max(simul$chl_a, na.rm  = T) &
           chl_a >= min(simul$chl_a, na.rm = T),
         tss <= max(simul$tss, na.rm  = T) &
           tss >= min(simul$tss, na.rm = T),
         doc <= max(simul$doc, na.rm  = T) &
           doc >= min(simul$doc, na.rm = T)) %>%
    mutate(type = ifelse(grepl('Lake',type),'Lake',type)) %>%
  expand_grid(.,chl_ratio = range) %>%
  mutate(power = ifelse(chl_ratio == 234, 0.57, 1),
         chl_a_biomass = exp(log(chl_ratio/1000)+log(chl_a)*power),
         tss_dead = tss-chl_a_biomass) %>%
  filter(chl_ratio == 234) %>%
  filter(tss_dead > 0.01)  


d234 <- no_sdd %>%
  mutate(kd_pred = predict(mod, .) + k_w,
         tss_kd = ((mod$coefficients[1])*tss_dead)/kd_pred,
         doc_kd = ((mod$coefficients[2])*doc)/kd_pred,
         chl_kd = ((mod$coefficients[3])*chl_a)/kd_pred,
         test_1 = tss_kd + doc_kd + chl_kd) 


```

# Results



```{r}

no_sech_tern <- ggtern(d234 , 
       aes(chl_kd,doc_kd,tss_kd,color = type)) + 
    geom_point(shape = 1, size = 0.5) +
    scale_color_manual(values = c('tan3','blue4','gray60')) + 
    theme_bw() + 
    theme_showarrows()  +
    labs(x = 'Algae', 
         y = 'DOC',
         z = 'TSS')  

no_sech_tern
ggsave(plot = no_sech_tern,
       filename = 'figs_2023/tern_aq.png',
       dpi=300,
       height = 6,
       width = 6)

```


### Ternaries by type with Jim's Hex zones

```{r}


# 
type_tern_sums <- d234 %>%
  group_by(type) %>%
  select(tss_kd,doc_kd,chl_kd,type) %>%
  summarize(across(where(is.numeric), mean))

# type_tern_sums
# 
# type_tern_sd <- d234 %>%
#   select(tss_kd,doc_kd,chl_kd,type) %>%
#   pivot_longer(tss_kd:chl_kd) %>%
#   group_by(type,name) %>%
#   summarize(across(where(is.numeric),
#                    list(mean = mean, 
#                         sd = sd),
#                    .names = '{.fn}')) %>%
#   mutate(mean1sd = mean + sd,
#          mean11sd = mean - sd,
#          mean2sd = mean + sd*2,
#          mean22sd = mean - sd*2)


# hex1 <- type_tern_sd %>%
#   select(mean1sd,type,name) %>%
#   pivot_wider(names_from = 'name',
#               values_from = 'mean1sd') %>%
#   bind_rows(type_tern_sd %>%
#   select(mean11sd,type,name) %>%
#   pivot_wider(names_from = 'name',
#               values_from = 'mean11sd'))
# 


tern_elip <- ggtern(d234,
       aes(chl_kd,doc_kd,tss_kd)) + 
    geom_point(shape = 1, size = 0.1,
               color = 'gray30') +
    scale_color_manual(values = c('tan3','blue4','gray60')) + 
    theme_bw() + 
    theme_showarrows()  +
    labs(x = 'Algae', 
         y = 'DOC',
         z = 'TSS')  + 
  facet_wrap(~type,
             nrow = 3) + 
  geom_mean_ellipse() + 
  #geom_hex_tern(bins = 50) + 
  geom_point(data = type_tern_sums, color = 'white') 




ggsave(plot = tern_elip,
       filename = 'figs_2023/tern_elip.png',
       dpi=300,
       height = 9,
       width = 4)

```




To have results we need to find some interesting candidate sites to explore. These
will be places that either 1) have lots of repeat data (temporal variation) or 2) 
have lots of spatial data (spatial variation), or ideally both!


## Temporal Variation


### Sites with long-term records

Long-term means:

  - At least 10 years of data with at least 3 observations in each year


```{r}

long_terms <- d234 %>%
  mutate(date = date(date_unity)) %>%
  group_by(lat,long,SiteID) %>%
  distinct(date,.keep_all = T) %>%
  mutate(year = year(date)) %>%
  mutate(n_years = n_distinct(year)) %>%
  filter(n_years > 5) %>%
  mutate(range = paste(min(year),max(year)))

# ggtern(d234, aes(tss_kd,doc_kd,chl_kd,
#                  size = secchi_pred)) + 
#     geom_point(shape = 19) +
#     scale_color_viridis_c() + 
#     theme_bw() + 
#     scale_size(range = c(0.1,3)) + 
#     theme_showarrows() + 
#     labs(x = 'TSS', 
#          y = 'DOC',
#          z = 'Algae') 

year_plotter <- function(df = nested_long$data[[1]]){
  ggtern(df, aes(tss_kd,doc_kd,chl_kd,color = year,
                 size = secchi_pred)) + 
    geom_point(shape = 19) +
    scale_color_viridis_c() + 
    theme_bw() + 
    scale_size(range = c(0.5,2)) + 
    theme_showarrows() + 
    labs(x = 'TSS', 
         y = 'DOC',
         z = 'Algae') + 
    ggtitle(paste(unique(df$id)))
}

nested_long <- long_terms %>%
  group_by(lat,long,SiteID, year, type) %>%
  mutate(n = n()) %>%
  filter(n > 2) %>%
  summarize(across(where(is.numeric), mean)) %>%
  group_by(lat, long, SiteID, type) %>%
  mutate(id = SiteID) %>%
  nest() %>%
  ungroup() %>%
  mutate(mean_plots = map(data, year_plotter))


site_sf <- long_terms %>%
  distinct(lat,long,SiteID,range, n_years) %>% 
  inner_join(nested_long) %>%
  st_as_sf(.,coords = c('lon','lat'), crs = 4326) %>%
  ungroup() 


# mapviewOptions(fgb = F)
# m1 <- mapview(site_sf,
#               zcol = 'n_years',
#               popup = leafpop::popupGraph(site_sf$mean_plots))
# 
# 
# 
# 
# mapshot(m1, url = 'long_terms.html')
```



### Seasonal variation


```{r}


season_sites <- d234 %>%
  filter(SiteID %in% site_sf$SiteID) %>%
  mutate(year = year(date_unity),
         month = month(date_unity)) %>%
  group_by(lat,long,SiteID,type,year) %>%
  summarize(dist_months = n_distinct(month)) %>%
  filter(dist_months > 5)
  



month_nest <- d234 %>%
  filter(SiteID %in% season_sites$SiteID) %>%
  mutate(year = year(date_unity),
         month = month(date_unity)) %>%
  group_by(SiteID, lat, long, month) %>%
  summarize(across(where(is.numeric),mean)) %>%
  group_by(SiteID, lat, lon) %>%
  mutate(id = SiteID) %>%
  nest()



month_plotter <- function(df =  month_nest$data[[1]]){
  ggtern(df, aes(tss_kd,doc_kd,chl_kd,color = month,
                        size = secchi_pred)) + 
    theme_bw() + 
    theme_showarrows() + 
    labs(x = 'TSS', 
         y = 'DOC',
         z = 'Alg') + 
    geom_point() + 
    scale_color_gradient2(low = 'green3',
                          mid = 'gray20',
                          high = 'purple3',
                          midpoint = 6) + 
    scale_size(range = c(0.5,3))
}

month_sf <- month_nest %>%
  st_as_sf(.,coords = c('lon','lat'), crs = 4326) %>%
  ungroup()  %>%
  mutate(month_plots = map(data, month_plotter)) 

# mapviewOptions(fgb = F)
# month_map <- mapview(month_sf,
#                      popup = leafpop::popupGraph(month_sf$month_plots))
# 
# mapshot(month_map, url = 'month_lon_terms.html')

```


## Type means


```{r}
month_site_means <- d234 %>%
  mutate(month = month(date_unity),
         year = year(date_unity),
         season = cut(month,4)) %>%
  group_by(lat, long, SiteID, season, type) %>%
  summarize(across(where(is.numeric), mean))


ggtern(month_site_means, aes(tss_kd,doc_kd,chl_kd,
                        size = secchi_pred)) + 
    theme_bw() + 
    theme_showarrows() + 
    labs(x = 'TSS', 
         y = 'DOC',
         z = 'Alg') + 
    geom_point() + 
    scale_color_gradient2(low = 'green3',
                          mid = 'gray20',
                          high = 'purple3',
                          midpoint = 6) + 
    scale_size(range = c(0.5,3)) + 
  facet_grid(season ~ type)



```



# MC/JJ Questions

- Exemplars for seasonal/long-term/temporal to show off ideas 

- Here's everything mapped over where we have tss/chl_a/doc and clustered?

- When making results drop secchi requirement and model secchi. 

- Huc membership

- Screening for long-term data at least 1/2 per year for > 20 years

- intra-annual at least 5 months per year. 

## Spatial Variation


```{r}


fox_ids <- tibble(SiteID = c('USGS-04084422','USGS-04084445','USGS-04084475','USGS-04085054',
             'USGS-04085059', 'USGS-04085139'),
             order = c(6,5,4,3,2,1))



fox_r <- d234  %>%
  inner_join(fox_ids) %>%
  mutate(date = as.Date(date_unity)) 


fox_r_map <- fox_r %>%
  distinct(SiteID, .keep_all = T) %>%
  st_as_sf(.,coords = c('lon','lat'),crs = 4326)

mapview(fox_r_map)

summary(fox_r)
cool_dates <- fox_r %>%
  group_by(date) %>%
  count() %>%
  arrange(-n) %>%
  ungroup() %>%
  slice(1:5) %>%
  arrange(date)



fox_days <- fox_r %>%
  dplyr::filter(date %in% cool_dates$date) 




ggtern(fox_days,aes(tss_kd,chl_kd,doc_kd, color = order)) + 
  geom_point() + 
  facet_wrap(~date,ncol=3) + 
  theme_bw() + 
  theme_showarrows() + 
  labs(x = 'TSS', 
       y = 'DOC',
       z = 'Alg') + 
  geom_point() + 
  geom_line() +
  scale_color_gradient(low = 'skyblue',high = 'blue3') + 
  theme(legend.position = c(0.8,0.2))


```


### Additions

- Geographic Coverage

- Plausible space (not deep clarity and high chl-a)

- Switch model application to wherever we have all data.

  - Requires a reasonable R2 to predict KD. Highly skewed in 
  the data
  
  - R2 round the corner. 



