---
title: "Ternary Light Partitioning"
author:
  - "Dr. Matthew Ross"
  - 'Dr. Jim Jawitz'
  - 'Dr. Matt Cohen'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::gitbook: default
  #bookdown::pdf_book: default
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 80
---

# Motivation and Introduction {.unnumbered}

Water clarity and its components (chlorophyll *a*, total suspended sediment
(TSS), and dissolved organic carbon (DOC)) are among the most functionally
important and socially-relevant aspects of water quality. Here we propose a
method for systematically evaluating the relative contribution of each water
clarity component to total clarity [[water column light extinction?]], a value
derived from Secchi disk depth. We further borrow a visualization tool from soil
analyses, the ternary diagram to describe these contributions: we build ternary
diagrams that partition light partitioning between DOC mass, derived algal
biomass, and derived non-algal mass.

[[add conceptual ternary diagram]]

## Workflow

[[is it intentional to use particle and mass/biomass interchangeably? Consider a
high-level consolidation of language below, I changed the above text to be
mass/biomass.]]

In its simplest form our workflow contains the following steps:

1.  Download all of [AquaSat](https://figshare.com/collections/AquaSat/4506140),
    a data set of four water quality parameters collated for the conterminous
    United States, and join the AquaSat pieces together to create a harmonized
    data set of the four parameters.

2.  Investigate the relationship between chlorohpyll *a*, TIS (inorganic
    sediment, derived from TSS), algal biomass (derived from chlorophyll *a*)
    and non-algal particles (derived from algal biomass and DOC). Create a
    dataset of algal biomass, non-algal particle mass, and dissolved organic
    carbon mass.

3.  Investigate data quality of AquaSat and select a subset of data where
    chlorophyll *a*, DOC, TSS and Secchi and create a dataset where all
    parameters were all measured simultaneously.

4.  Build a model that predicts light extinction using algal biomass, non-algal
    particles, and DOC where light extinction (K) is defined as:

    $$
    K = 1.4 / {Sechhi\:Disk\:Depth}
    $$

    [[note, the value I've seen for the constant is 1.7 (see Poole and Atkins
    1929), maybe cite here the associated paper for use of 1.4?]]

5.  Use this model to extract light extinction coefficients relative to each
    parameter (algal biomass, non-algal mass, DOC mass) and apply this model to
    visualizations in ternary space.

6.  Explore systematic controls on relative contribution to light extinction
    over space and time across CONUS.
    
    
```{r setup, include=FALSE}
# this code block checks for the packages used in this workflow, and installs 
# them if they are not installed on your machine.

package_installer <- function(x) {
  if (x %in% installed.packages()) {
    print(paste0('{', x ,'} package is already installed.'))
  } else {
    install.packages(x)
    print(paste0('{', x ,'} package has been installed.'))
  }
}

packages = c("tidyverse", "sf", "feather", "GGally", 
             "mapview", "bookdown", "ggthemes", "ggpmisc",
             "Metrics", "kableExtra", "broom")

lapply(packages, package_installer)
```


<!--chapter:end:index.Rmd-->

---
output: html_document
editor_options: 
  chunk_output_type: console
  markdown:
    wrap: 80
---


# (PART) Data downloads and decisions {-}

```{r library-ch1, include=FALSE}
library(sf)
library(feather)
library(GGally)
library(mapview)
library(ggthemes)
library(tidyverse)
```


# Download and Join


I recognize that I should probably just use `targets` for this part, but...
I (Matt Ross) don't know how to use targets yet! So, instead we are just
going to have several code chunks here that just downloads all the data directly
from figshare, preps data for analysis, etc...
and the chunks will just be set to `eval = FALSE`, so they won't
run unless you manually run it. That way this book can run in reasonable timeframes,
most other chunks will run based on "caching." Though, some data intensive tasks,
will also simply have `eval = F` to save runtime. These chunks will be flagged
explicitly, and all their data products saved to a `data/out` folder. 

## AquaSat download

AquaSat is held in a figshare collection with many pieces. For this work we
will need most of the data held there, which we will access through direct download.
The URLs are just from the AquaSat collection links.

```{r download-data, eval = F}
## AquaSat Download data. This takes > 2 hours,
## so only run if you think 
## AquaSat has been updated!

## Also

## In-situ raw data with methods
download.file('https://figshare.com/ndownloader/files/15475154',
              mode = 'wb',# Needs to be written in binary for some reason 
              destfile = 'data/in/aq_situ.zip')
unzip('data/in/aq_situ.zip',exdir = 'data/in/aq_situ')


# Site Inventory with type, because it's not in the other inventory
# Stupid aquasat developer (me/mattross!)
download.file('https://figshare.com/ndownloader/files/24720434',
              mode = 'wb',
              destfile = 'data/in/inv.feather')


## Unique site inventory 

download.file('https://figshare.com/ndownloader/files/24720437',
              mode = 'wb',
              destfile = 'data/in/unq_site.feather')

## Ecoregion data
download.file('https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/cec_na/na_cec_eco_l2.zip', destfile = 'data/in/eco2.zip')

unzip('data/in/eco2.zip',exdir = 'data/in/ecoregion')


```

## AquaSat Read and join. 

Note, also eval = F to save computation time

```{r read-join-aquasat, eval = F}


# This is only useful because it has type (lake/stream/etc...)
inv_type <- read_feather('data/in/inv.feather') %>%
    select(SiteID = MonitoringLocationIdentifier,
         type = ResolvedMonitoringLocationTypeName) %>%
  mutate(type = ifelse(grepl('Lake',type),'Lake',type))


#Real inventory
site.vis <- read_feather('data/in/unq_site.feather') %>%
  inner_join(inv_type) %>%
  distinct(SiteID,lat,long,type)



# Complete in-situ data, 6.5 million records. 
in.vis <- read_csv('data/in/aq_situ/in-situ/wqp_lagos_unity.csv') 

# Get clouds and path row sites

clouds <- read_feather('data/in/clouds.feather') %>%
  mutate(date = as.Date(SENSING_TIME)) %>%
  select(PATH = WRS_PATH, 
         ROW = WRS_ROW,
         clouds = CLOUD_COVER,
         date)


site_path_row <- site.vis %>%
  st_as_sf(.,coords = c('long','lat'), crs = 4326) %>%
  st_join(.,st_read('data/in/wrs/WRS2_descending.shp')) %>%
  select(SiteID, PATH, ROW) %>%
  as_tibble(.) %>%
  select(-geometry)

```

## Data Eval and Subset


### Selecting only simultaneous observations

I think the code mostly speaks for itself here
```{r simultaneous, eval = F}
# Only bother with data that has complete simultaneous observations of
# chl_a, doc, etc...

simul_vis <- in.vis %>%
  select(-p_sand) %>%
  filter(if_all(c(chl_a,doc,tss,secchi), ~!is.na(.))) %>% 
  inner_join(site.vis) %>%
  filter(type != 'Facility') %>%
  #Set some reasonable thresholds, AquaSat is too generous
  filter(secchi < 15,
         chl_a < 1000,## ug/L
         tss < 1000, ## mg/L
         doc < 50)

no_secchi <- in.vis %>%
  select(-p_sand) %>%
  filter(across(c(chl_a,doc,tss), ~!is.na(.))) %>% 
  inner_join(site.vis) %>%
  filter(type != 'Facility') %>%
  #Set some reasonable thresholds, AquaSat is too generous
  filter(
         chl_a < 1000,## ug/L
         tss < 1000, ## mg/L
         doc < 50)

# For speed
write_feather(simul_vis, 'data/out/simul.feather')
#For reproducibility
write_csv(simul_vis, 'data/out/simul.csv')

#For speed
write_feather(no_secchi, 'data/out/no_secchi.feather')



```



## Where are sites with simultaneous observations of clarity constituents? 


```{r map-simultaneous}

simul_vis <- read_csv('data/out/simul.csv')

unique_simul <- simul_vis %>%
  distinct(SiteID, lat, long, type) %>%
  st_as_sf(.,coords = c('long','lat'), crs = 4326)

mapviewOptions(fgb = FALSE) # remove flat geobuff which breaks display
mapview(unique_simul, zcol = 'type')

```


## What is the general relationiship between variables in log log space

```{r fig-four-vars-scatter, cache = T}


log_simul <- simul_vis %>%
  mutate(across(c(secchi,chl_a,tss,doc,tis),log10)) %>%
  filter(across(c(chl_a,doc,secchi,tss), ~!is.na(.) & . < Inf & . > -Inf)) 


log_simul %>%
  sample_frac(0.2) %>%
  ungroup() %>%
  select(secchi,chl_a,tss,doc,type) %>%
  ggpairs(lower = list(continuous = wrap('points',shape = 1)),
          diag = list(continuous = wrap('densityDiag', alpha = 0.5)),
         mapping = ggplot2::aes(color = type),
         columns = c('secchi','chl_a','tss','doc')) +
  theme_few() + 
  scale_color_manual(values = c('seagreen3','skyblue3','saddlebrown'))

```



<!--chapter:end:01_data_prep.Rmd-->

---
output: html_document
editor_options: 
  chunk_output_type: console
  markdown:
    wrap: 80
---

# Data quality assessments

Note, also eval = F to save computation time

```{r load-pkgs-ch2, include = F}
library(sf)
library(feather)
library(GGally)
library(mapview)
library(ggpmisc)
library(Metrics)
library(kableExtra)
library(broom)
library(ggthemes)
library(knitr)
library(tidyverse)

opts_chunk$set(warning = F, message = F)
```


## Chl-a to biomass

While it is more common in coastal and oceanic research, inland water scientists
rarely measure water clarity constituents in the way that is most directly 
transferable to our research question. Ideally we would routinely measure concentrations of
algae biomass, non-algal particles
(including inorganic sediment and organic sediment, I like to call this
stuff dead suspended sediment, but I guess NAP is fine), and dissolved organic 
carbon. These constituents all alter light penetration in water, but they are 
mutually exclusive groups with no measurement overlap between algal particles and NAP
for example. However, this is not what we usually measure. Instead, we have rich 
datasets of things that could indicate our desired constituents. These map as basically.

- Chlorophyll-a -> proxy for algae biomass.
- Total suspended solids (tss) -> All suspended solids, subtract algae biomass = NAP
- DOC -> DOC!

That means we have to basically do one key calculation. Convert chl-a to biomass.
Then subtract that from tss and assume that is NAP. While this is a absurdly
simple calculation it is filled with dangers. The relationship between chl-a
and algae biomass depends on temperature and nutrients and species and is 
[not universal](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2016GB005458).
Most studies focus on Chl:C ratios, but we are interested in Chl:Biomass ratio,
for which there are much fewer studies, though some studies below suggest 
a range of 0.005 to 0.1 Chl:biomass. (I need to find the C:Biomass study I used
forever ago.)

Papers that help:

  - https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lno.10338

  - https://pubs.usgs.gov/twri/twri9a7/twri9a7_7.4.pdf

  - https://www.researchgate.net/publication/230056019_Relationship_between_Chlorophyll-a_Concentration_and_Phytoplankton_Biomass_in_Several_Reservoirs_in_Czechoslovakia/link/5cb56c164585156cd79af804/download

  - https://archimer.ifremer.fr/doc/2005/publication-1172.pdf

### Approach

Our philosphy here is that we will use chl-a directly as a proxy for algal biomass
to minimize the number of transformations to this data, since we are primarily
interested in how algae alters light attenuation, regardless of how we estimate algae.
However, we still need to calculate Non-Algal Particle mass. For that we will
use a range of values of Chl:biomass and show the impact of picking a variety 
of ranges. 

*Caveats*

 - We don't have a true NAP to compare how accurate our approach is.
 
 - We do have Inorganic Sediment, but that is not the same thing as non-algal 
 particles, since it excludes particulate organic carbon. Still, where
 we have it, we will examine the relationship between NAP and TIS (total 
 inorganic sediment). 
 
 - Not entirely sure how we propogate this uncertainty downstream? Pick an average?



```{r data-nap-test}

in_vis <- read_feather('data/out/simul.feather')

no_sech <- read_feather('data/out/no_secchi.feather')

range <- c(50, 100, 200, 234)



#Couldn't think of a more clever way to multiply
#chl_a by the range of values, so just made
# dataframe 4 times bigger with new column called ratio. 
nap_test <- expand_grid(no_sech,chl_ratio = range) %>%
  mutate(power = ifelse(chl_ratio == 234, 0.57, 1),
         chl_a_biomass = exp(log(chl_ratio/1000)+log(chl_a)*power),
         tss_dead = tss-chl_a_biomass)


```

### Simple evaluations of conversion




#### Negative tss_dead

Probably the simplest way to see which ratios are more appropriate is to
look at how that impacts how many tss_dead observations are negative (impossible).

```{r neg-tss-dead}



nap_test %>%
  mutate(negative = ifelse(tss_dead < 0, 'negative', 'positive')) %>%
  group_by(negative,chl_ratio) %>%
  count() %>%
  pivot_wider(names_from = 'negative',values_from = 'n')  %>%
  mutate(percent_neg = negative/(positive+negative)*100) %>%
  kable()

```


Well, that sort of produces obvious results. Smaller ratios mean, less negative 
NAP estimates, but there is a big increase between 100-200, making me think the
"correct" ratio may be in there somewhere. 


#### Relationship between chl-a and tss_dead

If you recall from the previous section, there was a strong linear 
relationship between chl-a and tss across all sites and water types (r2 > 0.35).
One effect we'd expect to see with the tss_dead estimate is that this relationship
should decay. While there are reasons for chl-a to be correlated with tss (sediment
brings nutrients for example), chl-a also includes algal cells which are suspended
particles themselves. If this approach worked well, we should see weaker relationships
between tss_dead and chl-a


```{r fig-tss-dead-chla, cache = T, fig.width = 7,fig.height = 7, fig.cap = "At low chl_ratios (10/20), there is really no change to the correlation between tss and chl_a, but at higher ratios (100/200), the relationship does breakdown some, though not completely"}


#Subset for plotting purposes
nap_test %>%
  #Remove negatives and very small numbers (ug/L of sediment is basically zero)
  filter(tss_dead > 0.001) %>%
  sample_frac(0.1) %>%
ggplot(., aes(chl_a,tss_dead,color = type)) +
  facet_wrap(~chl_ratio) + 
  geom_point() + 
  scale_x_log10() + 
  scale_y_log10() + 
  stat_poly_eq() + 
  theme_few() + 
  scale_color_manual(values = c('seagreen3','skyblue3','saddlebrown')) 

```


- What to do with this? 
- 234*chl^0.57 seems logical but still high correlation in lakes between tss/chl_a,
but that may be a real thing for lakes? (corr is actually higher with this approach)



### Light attenuation model exploration

Model assumptions

- Not searching for intercept assuming observed 0.15 coefficient for intercept (pure water attenuation)

- For now, not doing the middle 2.5-97.5% of data, can add back in.

- Normality assumptions aren't fully met



```{r fig-kd-obs-pred, warning = F, cache = T, fig.cap = 'modelled light extinction coefficient versus observed, red line is 1:1 line and blue is line of best fit, R2 hardly varies between chl-a conversion approaches.'}


nap_est_no_secchi <- nap_test %>%
  ungroup() %>%
  filter(tss_dead > 0.01,
         secchi < 15 | is.na(secchi)) %>%
  mutate(secchi = ifelse(secchi < 0.1,0.1,secchi),
        kd = (1/(secchi)))



write_feather(nap_est_no_secchi %>%
                filter(chl_ratio == 234),'data/out/no_secchi_clean.feather')



nap_est <- nap_est_no_secchi %>%
  filter(!is.na(secchi))




k_w <- 1/1.5/max(nap_est$secchi)


kd_mod <- function(df){
  mod <- lm((kd-k_w)  ~ 0 + tss_dead + doc + chl_a, data = df)
}


kd_resid_extract <- function(df,mod){
  df <- df %>%
    mutate(residuals = mod$residuals,
           pred = mod$fitted.values + k_w)
}

# nap_est <- nap_est %>%
#   filter(SiteID %in% fine_sites$SiteID) 


nap_mods <- group_by(nap_est,chl_ratio) %>%
  nest() %>%
  mutate(mods = map(data,kd_mod),
         data = map2(data, mods, kd_resid_extract))

mod234 <- nap_mods %>%
  filter(chl_ratio == 234)



nap_resid <- nap_mods %>%
  select(-mods) %>%
  unnest(data) %>%
  mutate(ratio_rmse = paste('ratio =',chl_ratio,', R2 =', round(cor(pred,kd)^2,2)),
         year = year(date_unity))


nap_resid %>%
  sample_frac(0.2) %>%
  ggplot(., aes(kd,pred, color = year(date_unity))) + 
  geom_point(shape = 1) + 
  facet_wrap(~ratio_rmse) + 
  theme_few() + 
  xlab('kd (1/secchi)') +
  ylab('predicted kd') +
  scale_x_log10() + 
  scale_y_log10() +
  scale_color_viridis_c() + 
  stat_smooth(method = 'lm',se = F, color = 'black') +
  geom_abline(intercept = 0, slope = 1, col = 'red') 





```





### Variation in slopes w/ different chl biomass ratios

- Do all our painstaking decisions make a difference for model slopes?

- (They didn't impact model RMSE very much)

- No, they don't change slopes all that much either. 



```{r tab-nap-slope}

nap_slopes <- nap_mods %>%
  select(-data) %>%
  mutate(glance = map(mods, tidy)) %>%
  unnest(glance) %>%
  select(-mods,-std.error,-statistic) %>%
  pivot_wider(names_from = 'term', values_from = 'estimate')



kable(nap_slopes) %>%
  kable_paper() %>%
  scroll_box(width = '100%', height = '600px') %>%
  kable_styling(bootstrap_options = 'striped', full_width = )
```


# Final Decision

- Based on the above approaches and logic, the chl-a to biomass multiplier does 
not impact model performance very much. As such, we are going to use the 234 log 
ratio going forward. 




<!--chapter:end:02_data_qaqc.Rmd-->

